Wind veer and speed in turbulent Ekman flow part I: scaling analysis and universal profile model

### 2024-11-21 write1/manuscript.tex
ready to submit (?)

\begin{document} 
%
\maketitle
%
\begin{abstract}
  The profiles of wind speed and direction in turbulent Ekman flow
  are formulated based on asymptotic theory and data from direct numerical simulation. 
  %
  The profile of the streamwise component follows the classical viscous, logarithmic and wake scaling.
  %
  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
  boundary conditions that result in a reduction of the spiral-like rotation. 
  %
  The span-wise component poses a conceptual challenge to the channel-flow analogy
  in the context of asymptotic matching; it exhibits a mixed scaling in the surface layer, but follows
  outer scaling for most of the outer layer.
  %
  Viscous stress scales universally across the boundary layer in inner units while the total stress
  becomes universal as a function of outer height.
  %
  This implies a mixed scaling for the turbulent stress and eddy viscosity across the inner layer
  and convergence to a universal scaling as function of the outer height across the outer layer
  for increasing scale separation vide Reynolds numbers. 
\end{abstract}
~\\[1em] 
\textbf{Keywords} Direct Numerical Simulation $\cdot$ Scale Separation $\cdot$ Ekman Layer $\cdot$ Surface Layer $\cdot$ Hodograph

@@

    \includegraphics[width=\textwidth]{../plot/uv_innerouter.pdf}\\

@@

  \includegraphics[width=\textwidth]{../plot/stresses.pdf}

@@

  \includegraphics[width=\textwidth]{../plot/eddy_viscosity.pdf}\\

@@

    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}

@@

    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}

@@

  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}

@@

  \includegraphics[width=\textwidth]{../plot/ekman_ideal.pdf}

@@

\begin{figure}
  \begin{flushleft}
    \textbf{\ \ a)}\hspace{0.45\textwidth}\textbf{\ \ b)}\\
  \end{flushleft} 
  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
  \caption{
    \label{fig:profiles_outer}
    Shear-aligned velocity deficit for the streamwise (panel \textbf{(a)}) and spanwise (panel \textbf{(b)})
    components of the mean velocity $U^\alpha$ and $V^\alpha$.
    Solid lines show DNS data, dashed lines the Ekman profiles $U_\mathrm{ek}$ and $V_\mathrm{ek}$ as defined in
    Eq.~\ref{eqn:profile_ekman}. 
    %Region below $z^-=0.3$ is shown opaque. 
    Variations in $U_\mathrm{ek}$ and $V_\mathrm{ek}$ are
    a consequence of the normalization and related to changes in $u_\star$ and $\alpha_\star$ among the different $\RE_D$. 
  }
\end{figure}

@@

  \includegraphics[width=\textwidth]{../plot/u_profile.pdf}

@@

  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}

@@

  \centerline{\includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}}

@@

%
\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
  \caption{ \textbf{(a)} Velocity deficit, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
   In panels \textbf{(a)} and \textbf{(b)}, blue lines correspond to the streamwise component and red to the spanwise. }
  \label{fig:summary} 
\end{figure} 
The resulting velocity profiles across the entire boundary layer are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}. The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer. 
% 
The logarithmic scaling of the streamwise component is encountered as straight blue lines in panels \textbf{(a)} and \textbf{(b)} where 
the extent of the logarithmic range increases with $\RE$ towards lower values of $z^-$ and higher values of $z^+$ depending on the scaling. 
Importantly, the logarithmic region is widening for increasing Reynolds number -- irrespective of the scaling. 
% 
In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
% 
However, the velocity deficit in outer units becomes approximately universal, also across the inner layer; 
this reflects the compensation of reduced turning ($\alpha$) by increased drag ($u_\star$), 
and is consistent with the theoretical considerations discussed in \ref{sec:drag-law}. 
% 
\par
%
The spanwise velocity at a fixed height scales approximately as $Re_\tau^{-1}$ (Sec.~\ref{sec:spanwise}).
%
However, the fraction  of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer  (Fig.~\ref{fig:etling}a).  
%
This is because, in inner units, the inner layer grows as $Re_\tau$  which exactly compensates the reduced gradient of spanwise velocity. 
% 
The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter'  hodographs, corresponding to less turning, for increasing Reynolds number.

@@

\subsection{Comparison with other theories}
%%% VAN DRIEST SCALING
\begin{figure}
  \centerline{\includegraphics[width=0.66\textwidth]{../plot/vanDriest_profile.png}}
  \caption{Near-wall velocity profile according to the 
    van-Driest scaling (blue, solid) in comparison with 
    the present model (red, dashed), 
    the viscous law of the wall (orange dashed), and 
    the logarithmic law (green, solid)} 
  \label{fig:vanDriest}
\end{figure}  An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
where an exponential damping of Prandtl's mixing length is considered near the wall to yield 
\begin{align}
  \frac{\p u^+}{\p z^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa z^+\right)^2 (1-\exp\left[-z^+\right])}};
\end{align}
the spanwise component is zero as no rotational effects are considered. 
%
Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise velocity component (Fig.~\ref{fig:vanDriest}):
% 
Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$,  where the velocity transitions from the linear to the logarithmic profile. 
%
\par
%
%EKMAN SPIRAL 
For the higher layers of the ABL, the Ekman spiral is the simplest model available. 
%
When employed across the entirety of the ABL, the spiral of a turbulent Ekman layer is flattened with respect to Ekman's laminar solution, which  corresponds to a reduction of the veering angle both at the surface and throughout the ABL. 
%
We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).

@@

  \includegraphics[width=0.5\textwidth]{../plot/etling_direction_0.png} 
  \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_0.png}

@@

  \includegraphics[width=0.5\textwidth]{../plot/stress_rotation.pdf}
  \includegraphics[width=0.5\textwidth]{../plot/misalignment.pdf} 

@@

%
%\begin{figure}
%  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
%  \caption{Total velocity and veering (in degrees) vs inner and outer height.
%    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
%\end{figure} 
%
\section{Conclusions}
%
We investigate the wind veer and fundamental scaling properties of the velocity profiles in Ekman flow. 
%
Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.  
% 
This formulation is consistent with the DNS data and also yields reasonable results  at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
%
The classic formulation of the Ekman layer, employing the surface boundary condition, is replaced by a modified solution that can be obtained by Ekman's system of governing equations, but using different boundary conditions that are more appropriate of the actual situation encountered in the planetary boundary layer. 
% 
The three parameters that characterize this boundary condition are estimated based on DNS data. 
% 
\par
% 
To quantify the spanwise velocity component consistently across the boundary layer, we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
%
\begin{equation} 
  V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+)
\end{equation}
%
that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
% 
This scaling is derived here based on scaling considerations, and it is in excellent agreement with the DNS data available.   
% 
In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model discussed above (Sec.~\ref{sec:ekman}).
% 
\par
% 
Our results suggest that there is no lower limit of the turning. 
% 
Hence---despite its very large scale separation / huge Reynolds number---the ABL is not 
in the 'limit' of high Reynolds number from the perspective of wind veer, but always in a regime where 
changes in $\RE$ impact on the vertical partitioning of rotation.

@@

\vspace{1em}
\noindent\textbf{Acknowledgements} This study is funded by the European Commission's 7$^{th}$ Framework Programme through 
the ERC 2019 Starting Grant trainABL (ID 851374). Simulations were carried out on the German supercomputing systems 
\texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
For\-schungs\-zen\-trum J\"ulich (Projects \texttt{hku24} and \texttt{stadit}). 
We extend our thanks to Sally Issa for valuable feedback on the manuscript.

@@

\noindent\textbf{Funding} CA is funded by ERC Starting Grant \emph{trainABL} (ID 851374). Open Access funding for this publication through the DEAL agreement. 
\vspace{1em}

@@

\bibliographystyle{spbasic_updated}  
\bibliography{ansorge}

>>>>

\begin{document} 
%
\maketitle
%
\begin{abstract}
  The profiles of wind speed and direction in turbulent Ekman flow
  are formulated based on asymptotic theory and data from direct numerical simulation. 
  %
  The profile of the streamwise component follows the classical viscous, logarithmic and wake scaling.
  %
  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
  boundary conditions that result in a reduction of the spiral-like rotation. 
  %
  The span-wise component poses a conceptual challenge to the channel-flow analogy
  in the context of asymptotic matching; it exhibits a mixed scaling in the surface layer, but follows
  outer scaling for most of the outer layer.
  %
  Viscous stress scales universally across the boundary layer in inner units while the total stress
  becomes universal as a function of outer height.
  %
  This implies a mixed scaling for the turbulent stress and eddy viscosity across the inner layer
  and convergence to a universal scaling as function of the outer height across the outer layer
  for increasing scale separation vide Reynolds numbers. 
  % 
  The extrapolation to these scaling to atmospheric scale separation is confirmed via large-eddy simulation 
  in part II of this manuscript. 
\end{abstract}
~\\[1em] 
\textbf{Keywords} Direct Numerical Simulation $\cdot$ Scale Separation $\cdot$ Ekman Layer $\cdot$ Surface Layer $\cdot$ Hodograph

@@

    \includegraphics[width=\textwidth]{Fig1.eps}\\ %uv_innerouter.eps}\\

@@

  \includegraphics[width=\textwidth]{Fig2ab.eps}\\  %stresses.eps}

@@

  \includegraphics[width=\textwidth]{Fig2cd.eps}\\ %-eddy_viscosity.eps}\\

@@

    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{Fig3a.png}
    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{Fig3b.png}
    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{Fig3c.png}}

@@

    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{Fig3d.png}}

@@

  \centerline{\includegraphics[width=0.7\textwidth]{Fig4.eps}} % ustar_alpha.eps}}

@@

  \includegraphics[width=\textwidth]{Fig5.eps} %ekman_ideal.eps}

@@

\begin{figure}
  \begin{flushleft}
    \textbf{\ \ a)}\hspace{0.45\textwidth}\textbf{\ \ b)}\\
  \end{flushleft} 
  \includegraphics[width=0.5\textwidth]{Fig6a.eps} %outer_layer_u.eps}
  \includegraphics[width=0.5\textwidth]{Fig6b.eps} %outer_layer_w.eps}
  \caption{
    \label{fig:profiles_outer}
    Shear-aligned velocity deficit for the streamwise (panel \textbf{(a)}) and spanwise (panel \textbf{(b)})
    components of the mean velocity $U^\alpha$ and $V^\alpha$.
    Solid lines show DNS data, dashed lines the Ekman profiles $U_\mathrm{ek}$ and $V_\mathrm{ek}$ as defined in
    Eq.~\ref{eqn:profile_ekman}. 
    %Region below $z^-=0.3$ is shown opaque. 
    Variations in $U_\mathrm{ek}$ and $V_\mathrm{ek}$ are
    a consequence of the normalization and related to changes in $u_\star$ and $\alpha_\star$ among the different $\RE_D$. 
  }
\end{figure}

@@

  \includegraphics[width=\textwidth]{Fig7.eps} % u_profile.eps}

@@

  \includegraphics[width=1.0\textwidth]{Fig8.eps} %w_profile.eps}

@@

  \centerline{\includegraphics[width=0.5\textwidth]{Fig9.eps}} %c_log.eps}}

@@

%
\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{Fig10.eps}  %ekman_profiles.eps 
  \caption{ \textbf{(a)} Velocity deficit, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
   In panels \textbf{(a)} and \textbf{(b)}, blue lines correspond to the streamwise component and red to the spanwise. }
  \label{fig:summary} 
\end{figure} 
The resulting velocity profiles across the entire boundary layer are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}. The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer. 
% 
The logarithmic scaling of the streamwise component is encountered as straight blue lines in panels \textbf{(a)} and \textbf{(b)} where 
the extent of the logarithmic range increases with $\RE$ towards lower values of $z^-$ and higher values of $z^+$ depending on the scaling. 
Importantly, the logarithmic region is widening for increasing Reynolds number -- irrespective of the scaling. 
% 
In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
% 
However, the velocity deficit in outer units becomes approximately universal, also across the inner layer; 
this reflects the compensation of reduced turning ($\alpha$) by increased drag ($u_\star$), 
and is consistent with the theoretical considerations discussed in \ref{sec:drag-law}. 
% 
\par
%
The spanwise velocity at a fixed height scales approximately as $Re_\tau^{-1}$ (Sec.~\ref{sec:spanwise}).
%
However, the fraction  of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer  (Fig.~\ref{fig:etling}a).  
%
This is because, in inner units, the inner layer grows as $Re_\tau$  which exactly compensates the reduced gradient of spanwise velocity. 
% 
The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter'  hodographs, corresponding to less turning, for increasing Reynolds number.

@@

\subsection{Comparison with other theories}
%%% VAN DRIEST SCALING
\begin{figure}
  \centerline{\includegraphics[width=0.66\textwidth]{Fig11.eps}} %vanDriest_profile.eps}}
  \caption{Near-wall velocity profile according to the 
    van-Driest scaling (blue, solid) in comparison with 
    the present model (red, dashed), 
    the viscous law of the wall (orange dashed), and 
    the logarithmic law (green, solid)} 
  \label{fig:vanDriest}
\end{figure}  An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
where an exponential damping of Prandtl's mixing length is considered near the wall to yield 
\begin{align}
  \frac{\p u^+}{\p z^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa z^+\right)^2 (1-\exp\left[-z^+\right])}};
\end{align}
the spanwise component is zero as no rotational effects are considered. 
%
Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise velocity component (Fig.~\ref{fig:vanDriest}):
% 
Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$,  where the velocity transitions from the linear to the logarithmic profile. 
%
\par
%
%EKMAN SPIRAL 
For the higher layers of the ABL, the Ekman spiral is the simplest model available. 
%
When employed across the entirety of the ABL, the spiral of a turbulent Ekman layer is flattened with respect to Ekman's laminar solution, which  corresponds to a reduction of the veering angle both at the surface and throughout the ABL. 
%
We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).

@@

  \includegraphics[width=0.5\textwidth]{Fig12a.eps} %etling_direction.eps} 
  \includegraphics[width=0.5\textwidth]{Fig12b.eps} %etling_magnitude.eps}

@@

  \includegraphics[width=0.5\textwidth]{Fig13a.eps} %stress_rotation.eps}
  \includegraphics[width=0.5\textwidth]{Fig13b.eps} %misalignment.eps} 

@@

%
%\begin{figure}
%  \includegraphics[width=\textwidth]{mag_dir.eps} 
%  \caption{Total velocity and veering (in degrees) vs inner and outer height.
%    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
%\end{figure} 
%
\section{Conclusions}
%
We investigate the wind veer and fundamental scaling properties of the velocity profiles in Ekman flow. 
%
Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.  
% 
This formulation is consistent with the DNS data and also yields reasonable results  at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
%
The classic formulation of the Ekman layer, employing the surface boundary condition, is replaced by a modified solution that can be obtained by Ekman's system of governing equations, but using different boundary conditions that are more appropriate of the actual situation encountered in the planetary boundary layer. 
% 
The three parameters that characterize this boundary condition are estimated based on DNS data. 
% 
\par
% 
To quantify the spanwise velocity component consistently across the boundary layer, we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
%
\begin{equation} 
  V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+)
\end{equation}
%
that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
% 
This scaling is derived here based on scaling considerations, and it is in excellent agreement with the DNS data available.   
% 
In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model discussed above (Sec.~\ref{sec:ekman}).
% 
\par
% 
Our results suggest that there is no lower limit of the turning. 
% 
Hence---despite its very large scale separation / huge Reynolds number---the ABL is not 
in the 'limit' of high Reynolds number from the perspective of wind veer, but always in a regime where 
changes in $\RE$ impact on the vertical partitioning of rotation.

@@

\vspace{1em}
\noindent\textbf{Acknowledgements} Simulations were carried out on the German supercomputing systems 
\texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
For\-schungs\-zen\-trum J\"ulich (Projects \texttt{hku24} and \texttt{stadit}). 
We extend our thanks to Sally Issa for valuable feedback on the manuscript.

@@

\noindent\textbf{Funding} CA is funded by the European Commission's 7$^{th}$ Framework Programme through 
the ERC 2019 Starting Grant trainABL (ID 851374). HW is funded by the Ministry for Science and Culture of 
Lower Saxony (project \emph{Zukunftskonzept Windenergieforschung}). Open Access funding for this publication through the DEAL agreement. 
\vspace{1em}

@@

\bibliographystyle{spbasic_updated}  
% \bibliography{ansorge}
\begin{thebibliography}{57}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Ansorge(2019)}]{ansorge:BM2019}
Ansorge C (2019) Scale {{Dependence}} of {{Atmosphere}}--{{Surface Coupling
  Through Similarity Theory}}. Boundary-Layer Meteorol 170(1):1--27,
  \doi{10.1007/s10546-018-0386-y}

\bibitem[{Ansorge(2024{\natexlab{a}})}]{ansorge:2024c}
Ansorge C (2024{\natexlab{a}}) Direct numerical simulation of turbulent
  {{Ekman}} flow ({{Re}}=0500): {{DOI}} 10.17169/refubium-42505.
  \doi{10.17169/refubium-42505}

\bibitem[{Ansorge(2024{\natexlab{b}})}]{ansorge:2024a}
Ansorge C (2024{\natexlab{b}}) Direct numerical simulation of turbulent
  {{Ekman}} flow ({{Re}}=1000): {{DOI}} 10.17169/refubium-42507.
  \doi{10.17169/refubium-42507}

\bibitem[{Ansorge(2024{\natexlab{c}})}]{ansorge:2024}
Ansorge C (2024{\natexlab{c}}) Direct numerical simulation of turbulent
  {{Ekman}} flow ({{Re}}=1300): {{DOI}} 10.17169/refubium-42508.
  \doi{10.17169/refubium-42508}

\bibitem[{Ansorge(2024{\natexlab{d}})}]{ansorge:2024b}
Ansorge C (2024{\natexlab{d}}) Direct numerical simulation of turbulent
  {{Ekman}} flow ({{Re}}=1600); {{DOI}} 10.17169/refubium-42509.
  \doi{10.17169/refubium-42509}

\bibitem[{Ansorge and Mellado(2014)}]{ansorge:BM2014}
Ansorge C, Mellado JP (2014) Global {{Intermittency}} and {{Collapsing
  Turbulence}} in the {{Stratified Planetary Boundary Layer}}. Boundary-Layer
  Meteorol 153(1):89--116, \doi{10.1007/s10546-014-9941-3}

\bibitem[{Ansorge and Mellado(2016)}]{ansorge:JFM2016}
Ansorge C, Mellado JP (2016) Analyses of external and global intermittency in
  the logarithmic layer of {{Ekman}} flow. J Fluid Mech 805:611--635,
  \doi{10.1017/jfm.2016.534}

\bibitem[{Baars and Marusic(2020)}]{baars:JFM2020a}
Baars WJ, Marusic I (2020) Data-driven decomposition of the streamwise
  turbulence kinetic energy in boundary layers. {{Part}} 2. {{Integrated}}
  energy and. J Fluid Mech 882:A26, \doi{10.1017/jfm.2019.835}

\bibitem[{Barenblatt(1993)}]{barenblatt:JFM1993}
Barenblatt GI (1993) Scaling laws for fully developed turbulent shear flows.
  {{Part}} 1. {{Basic}} hypotheses and analysis. J Fluid Mech 248:513--520,
  \doi{10.1017/S0022112093000874}

\bibitem[{Barenblatt and Goldenfeld(1995)}]{barenblatt:PF1995}
Barenblatt GI, Goldenfeld N (1995) Does fully developed turbulence exist?
  {{Reynolds}} number independence versus asymptotic covariance. Phys Fluids
  7(12):3078--3082, \doi{10/bw2sqs}

\bibitem[{Blackadar and Tennekes(1968)}]{blackadar:JAS1968}
Blackadar AK, Tennekes H (1968) Asymptotic {{Similarity}} in {{Neutral
  Barotropic Planetary Boundary Layers}}. Journal of the Atmospheric Sciences
  25:1015--1020, \doi{10.1175/1520-0469(1968)025<1015:ASINBP>2.0.CO;2}

\bibitem[{Brown et~al.(2005)Brown, Beljaars, Hersbach, Hollingsworth, Miller,
  and Vasiljevic}]{brown:QJR2005}
Brown AR, Beljaars ACM, Hersbach H, Hollingsworth A, Miller M, Vasiljevic D
  (2005) Wind turning across the marine atmospheric boundary layer. Quarterly
  Journal of the Royal Meteorological Society 131(607):1233--1250,
  \doi{10/dw2bxz}

\bibitem[{Calaf et~al.(2010)Calaf, Meneveau, and Meyers}]{calaf:PF2010}
Calaf M, Meneveau C, Meyers J (2010) Large eddy simulation study of fully
  developed wind-turbine array boundary layers. Phys Fluids 22(1):015,110,
  \doi{10/b7gc6v}

\bibitem[{Coleman et~al.(1992)Coleman, Ferziger, and Spalart}]{coleman:JFM1992}
Coleman GN, Ferziger JH, Spalart PR (1992) Direct {{Simulation}} of the
  {{Stably Stratified Turbulent Ekman Layer}}. Journal of Fluid Mechanics
  244:677--712, \doi{10.1017/S0022112092003264}

\bibitem[{{da Silva} et~al.(2014){da Silva}, Hunt, Eames, and
  Westerweel}]{dasilva:ARF2014}
{da Silva} CB, Hunt JC, Eames I, Westerweel J (2014) Interfacial {{Layers
  Between Regions}} of {{Different Turbulence Intensity}}. Annu Rev Fluid Mech
  46(1):567--590, \doi{10.1146/annurev-fluid-010313-141357}

\bibitem[{Dimotakis(2005)}]{dimotakis:ARF2005}
Dimotakis PE (2005) {{TURBULENT MIXING}}. Annual Review of Fluid Mechanics
  37(1):329--356, \doi{10.1146/annurev.fluid.36.050802.122015}

\bibitem[{Ekman(1905)}]{ekman:AMA1905}
Ekman VW (1905) On the influence of the earth's rotation on ocean currents. Ark
  Mat Astron Fys, Vol 2 (1905), pp 1-53 2:1--53

\bibitem[{Ellison(1955)}]{ellison:QJR1955}
Ellison TH (1955) The {{Ekman}} spiral. Q J Roy Met Soc 81(350):637--638,
  \doi{10.1002/qj.49708135025}

\bibitem[{Emeis(2018)}]{emeis:2018}
Emeis S (2018) Wind {{Energy Meteorology}}, 2nd edn. Atmospheric {{Physics}}
  for {{Wind Power Generation}}, Springer, Heidelberg

\bibitem[{Emeis et~al.(2007)Emeis, {Baumann-Stanzer}, Piringer, Kallistratova,
  Kouznetsov, and Yushkov}]{emeis:m2007}
Emeis S, {Baumann-Stanzer} K, Piringer M, Kallistratova M, Kouznetsov R,
  Yushkov V (2007) Wind and turbulence in the urban boundary layer analysis
  from acoustic remote sensing data and fit to analytical relations. metz
  16(4):393--406, \doi{10.1127/0941-2948/2007/0217}

\bibitem[{Esau(2004)}]{esau:EFM2004}
Esau I (2004) Simulation of {{Ekman Boundary Layers}} by {{Large Eddy Model}}
  with {{Dynamic Mixed Subfilter Closure}}. Environmental Fluid Mechanics
  4(3):273--303, \doi{10/b8f3kh}

\bibitem[{Etling(2002)}]{etling:2002}
Etling D (2002) Theoretische {{Meteorologie}}, 2nd edn. Eine
  {{Einf{\"u}hrung}}, Springer-Verlag, Berlin, Heidelberg

\bibitem[{Etling(2008)}]{etling:2008}
Etling D (2008) {Theoretische Meteorologie: eine Einf{\"u}hrung}, 3rd edn.
  Springer, Berlin Heidelberg

\bibitem[{Foken(2002)}]{foken:m2002}
Foken T (2002) Some aspects of the viscous sublayer. metz 11(4):267--272,
  \doi{10.1127/0941-2948/2002/0011-0267}

\bibitem[{Foken et~al.(1978)Foken, Kitajgorodskij, and Kuznecov}]{foken:BM1978}
Foken {\relax Th}, Kitajgorodskij SA, Kuznecov OA (1978) On the dynamics of the
  molecular temperature boundary layer above the sea. Boundary-Layer Meteorol
  15(3):289--300, \doi{10.1007/BF02652602}

\bibitem[{Ghannam and {Bou-Zeid}(2021)}]{ghannam:QJR2021}
Ghannam K, {Bou-Zeid} E (2021) Baroclinicity and directional shear explain
  departures from the logarithmic wind profile. Quarterly Journal of the Royal
  Meteorological Society 147:434--464, \doi{10/gnj6z2}

\bibitem[{Gryning et~al.(2007)Gryning, Batchvarova, Br{\"u}mmer, J{\o}rgensen,
  and Larsen}]{gryning:BM2007}
Gryning SE, Batchvarova E, Br{\"u}mmer B, J{\o}rgensen H, Larsen S (2007) On
  the extension of the wind profile over homogeneous terrain beyond the surface
  boundary layer. Boundary-Layer Meteorol 124(2):251--268,
  \doi{10.1007/s10546-007-9166-9}

\bibitem[{H{\"o}gstr{\"o}m(1988)}]{hogstrom:BM1988}
H{\"o}gstr{\"o}m U (1988) Non-dimensional wind and temperature profiles in the
  atmospheric surface layer: {{A}} re-evaluation. Boundary-Layer Meteorology
  42:55--78, \doi{10.1007/BF00119875}

\bibitem[{H{\"o}gstr{\"o}m(1996)}]{hogstrom:BM1996}
H{\"o}gstr{\"o}m U (1996) Review of some basic characteristics of the
  atmospheric surface layer. Boundary-Layer Meteorology 78(3-4):215--246,
  \doi{10.1007/BF00120937}

\bibitem[{Jacobs and Van~Boxel(1988)}]{jacobs:AFM1988}
Jacobs AF, Van~Boxel JH (1988) Changes of the displacement height and roughness
  length of maize during a growing season. Agricultural Forest Meteorol
  42(1):53--62, \doi{10.1016/0168-1923(88)90066-4}

\bibitem[{Jiang et~al.(2018)Jiang, Wang, and Sullivan}]{jiang:JAS2018}
Jiang Q, Wang S, Sullivan P (2018) Large-{{Eddy Simulation Study}} of {{Log
  Laws}} in a {{Neutral Ekman Boundary Layer}}. Journal of the Atmospheric
  Sciences 75(6):1873--1889, \doi{10.1175/JAS-D-17-0153.1}

\bibitem[{Jim{\'e}nez(2012)}]{jimenez:ARF2012}
Jim{\'e}nez J (2012) Cascades in {{Wall-Bounded Turbulence}}. Ann Rev Fluid
  Mech 44(Volume 44, 2012):27--45, \doi{10.1146/annurev-fluid-120710-101039}

\bibitem[{Kelly and Gryning(2010)}]{kelly:BM2010}
Kelly M, Gryning SE (2010) Long-{{Term Mean Wind Profiles Based}} on
  {{Similarity Theory}}. Boundary-Layer Meteorol 136(3):377--390,
  \doi{10.1007/s10546-010-9509-9}

\bibitem[{Kelly and Troen(2016)}]{kelly:WE2016}
Kelly M, Troen I (2016) Probabilistic stability and `tall' wind profiles:
  Theory and method for use in wind resource assessment. Wind Energy
  19(2):227--241, \doi{10.1002/we.1829}

\bibitem[{Klein et~al.(2021)Klein, Maier, and Schmidt}]{klein:PAM2021a}
Klein M, Maier RE, Schmidt H (2021) Stochastic modeling of transient neutral
  and stably-stratified {{Ekman}} boundary layers. P A M M 21(1):e202100,146,
  \doi{10.1002/pamm.202100146}

\bibitem[{Li et~al.(2018)Li, Gentine, Mellado, and McColl}]{li:JAS2018}
Li Q, Gentine P, Mellado JP, McColl KA (2018) Implications of {{Nonlocal
  Transport}} and {{Conditionally Averaged Statistics}} on {{Monin}}--{{Obukhov
  Similarity Theory}} and {{Townsend}}'s {{Attached Eddy Hypothesis}}. J Atmos
  Sci 75(10):3403--3431, \doi{10.1175/JAS-D-17-0301.1}

\bibitem[{Lindvall and Svensson(2019)}]{lindvall:QJR2019}
Lindvall J, Svensson G (2019) Wind turning in the atmospheric boundary layer
  over land. Q J Roy Met Soc 145(724):3074--3088, \doi{10.1002/qj.3605}

\bibitem[{Mellado and Ansorge(2012)}]{mellado:ZaM2012}
Mellado J, Ansorge C (2012) Factorization of the {{Fourier}} transform of the
  pressure-{{Poisson}} equation using finite differences in colocated grids. Z
  angew Math Mech 92(5):380--392, \doi{10.1002/zamm.201100078}

\bibitem[{Mellado et~al.(2016)Mellado, {van Heerwaarden}, and
  Garcia}]{mellado:BM2016}
Mellado JP, {van Heerwaarden} CC, Garcia JR (2016) Near-{{Surface Effects}} of
  {{Free Atmosphere Stratification}} in {{Free Convection}}. Boundary-Layer
  Meteorol 159(1):69--95, \doi{10.1007/s10546-015-0105-x}
